### 2020-11-05 latex/acrobot.tex
Updated a lot of the proof, almost done oscillations for acrobot

Let us now compute the constrained dynamics under
(\ref{eqn:acrobot-constraint}).
Note that \(dh_q = \begin{bmatrix}0 & 1\end{bmatrix}\), while
\[
    dh_{p_u} = \frac{-\bar{q}_a I}{1 + I^2 p_u^2}
    .
\]
Inserting these to (\ref{eqn:g-qupu}), we get the solution for \(p_a\) on the
constraint manifold:
\[
    p_a(q_u,p_u) = \frac{
        (1+c_a)(1+I^2 p_u^2)p_u - m^2gl^3\bar{q}_a I (2-c_a^2)(2s_u + s_{ua})
    }{ml^2(3+2c_a)(1+I^2 p_u^2)}
    .
\]
The dynamics for \(p_u\) do not contain \(p_a\), so they remain unchanged.
The constrained dynamics for \(q_u\) are given by 
\begin{equation*}
    \dot{q}_u = e_1\tpose \Minv(q) \begin{bmatrix}
                    p_u \\ p_a(q_u,p_u)
                \end{bmatrix} %\\
    ,
\end{equation*}
%        &= \frac{1}{ml^2(2-c_a^2)} 
%            \left( p_u - 
%                (1+c_a)\frac{(1+c_a)(1+I^2 p_u^2)p_u - m^2gl^3\bar{q}_a I (2-c_a^2)(2s_u + s_{ua})
%                }{ml^2(3+2c_a)(1+I^2 p_u^2)}
%            \right) 
%        ,
%\end{align*}
which can be simplified to 
\begin{equation*}
    \dot{q_u} = \frac{(1+I^2 p_u^2)p_u - m^2gl^3\bar{q}_a I(2s_u + s_{ua})(1+c_a) }{ml^2(1+I^2 p_u^2)(3+2c_a)}
    .
\end{equation*}
Hence, the constrained dynamics for the acrobot under
(\ref{eqn:acrobot-constraint}) are
\begin{equation}\label{eqn:acrobot-constrained-dynamics}
    \begin{cases}
        \dot{q}_u = \frac{(1+I^2 p_u^2)p_u - m^2gl^3\bar{q}_a I(2s_u + s_{ua})(1+c_a) }
            {ml^2(1+I^2 p_u^2)(3+2c_a)}
        , \\
        \dot{p}_u = - m g l (2s_u + s_{ua})
        .
    \end{cases}
\end{equation}

@@

Letting \(I = 0\) we get the nominal system
\begin{equation}\label{eqn:khalil-perturbation-nominal}
    \begin{cases}
        \dot{x}_0 = f(t,x,0) ,\\
        x_0(t_0) = \eta_0 ,
    \end{cases}
\end{equation}
which we need to solve for the explicit solution \(x_0(t)\) on \([t_0,t_1]\).
One can use \(x_0(t)\) to compute a first-order approximation of
the nonlinear system
\begin{equation}\label{eqn:khalil-perturbation-firstorder}
    \begin{cases}
        \dot{x}_1 = \pdiff{f}{x}(t,x_0(t),0)x_1 + \pdiff{f}{I}(t,x_0(t),0)
        , \\
        x_1(t_0) = 0, \\
    \end{cases}
\end{equation}
which is a linear time-varying system with solution \(x_1(t)\).

We now paraphrase Khalil's Theorem 10.1 \cite{khalil_nonlinear} on the
accuracy of perturbation analysis.
\begin{thm}
    Suppose \(f : [t_0,t_1] \times D \times [-I_0,I_0] \rightarrow \R^2\) is
    \(C^1\), and that the nominal problem (\ref{eqn:khalil-perturbation-nominal}) has a
    unique solution \(x_0(t) \in D\) on \([t_0,t_1]\).
    Then there exists \(I^\star > 0\) such that for all \(|I| < I^\star\), the
    solution \(x(t,I)\) to (\ref{eqn:khalil-setup}) satisfies
    \[
        \norm{x(t,I) - \left(x_0(t) + I x_1(t)\right)} \leq k |I^2|
    \]
    for some \(k > 0\).
\end{thm}

The above Theorem tells us that we can approximate the solution of the nonlinear
system by the Taylor approximation \(x_0(t) + I x_1(t)\).
In other words, when \(I\) is small enough solutions of the nonlinear system and
the Taylor approximation are the same up to order \(I^2\) along compact time
intervals.
Since we know that the acrobot behaves like a pendulum at \(I = 0\), we can
use this theory to more easily study the behaviour of orbits and prove energy
injection properties of our VNHC.

@@

The acrobot's constrained dynamics in
\((\alpha,\mu)\)-coordinates can be computed by evaluating
\[
    \left.\begin{cases}
        \dot{\mu} = 
        \begin{bmatrix} \pdiff{\mu}{q_u} & \pdiff{\mu}{p_u} \end{bmatrix}
        \begin{bmatrix} \dot{q}_u \\ \dot{p}_u \end{bmatrix}
        , \\
        \dot{\alpha} = \begin{bmatrix} \pdiff{\alpha}{q_u} & \pdiff{\alpha}{p_u} \end{bmatrix}
        \begin{bmatrix} \dot{q}_u \\ \dot{p}_u \end{bmatrix}
        .
    \end{cases} \right|_{(q_u,p_u) = T\inv(\mu,\alpha)}
\]
These dynamics are incredibly complicated and difficult to express, so we will
write them simply as
\[
    \begin{cases}
        \dot{\mu} = f_\mu(\alpha,\mu,I)
        ,\\
        \dot{\alpha} = f_\alpha(\alpha,\mu,I)
        .
    \end{cases}
\]

To find the nominal system we set \(I = 0\); 
MATLAB's symbolic toolbox operations tell us that
\begin{equation}\label{eqn:acrobot-nominal-ma-dynamics}
    \begin{cases}
        \dot{\mu} = 0
        , \\
        \dot{alpha} = \sqrt{\frac{6g}{5l}} 
        \sqrt{\frac{\cos(\mu\cos(\alpha)) - \cos(\mu)}{\mu^2 \sin(\alpha)^2}}
        .
    \end{cases}
\end{equation}

>>>>

Let us now compute the constrained dynamics under
(\ref{eqn:acrobot-constraint}).
Note that \(dh_q = \begin{bmatrix}0 & 1\end{bmatrix}\), while
\[
    dh_{p_u} = \frac{-\bar{q}_a I}{1 + I^2 p_u^2}
    .
\]
Inserting these to (\ref{eqn:g-qupu}), we get the solution for \(p_a\) on the
constraint manifold:
\[
    p_a(q_u,p_u) = \frac{
        (1+c_a)(1+I^2 p_u^2)p_u - m^2gl^3\bar{q}_a I (2-c_a^2)(2s_u + s_{ua})
    }{ml^2(3+2c_a)(1+I^2 p_u^2)}
    .
\]
The dynamics for \(p_u\) do not contain \(p_a\), so they remain unchanged.
The constrained dynamics for \(q_u\) are given by 
\begin{equation*}
    \dot{q}_u = e_1\tpose \Minv(q) \begin{bmatrix}
                    p_u \\ p_a(q_u,p_u)
                \end{bmatrix} %\\
    ,
\end{equation*}
%        &= \frac{1}{ml^2(2-c_a^2)} 
%            \left( p_u - 
%                (1+c_a)\frac{(1+c_a)(1+I^2 p_u^2)p_u - m^2gl^3\bar{q}_a I (2-c_a^2)(2s_u + s_{ua})
%                }{ml^2(3+2c_a)(1+I^2 p_u^2)}
%            \right) 
%        ,
%\end{align*}
which can be simplified to 
\begin{equation*}
    \dot{q_u} = \frac{(1+I^2 p_u^2)p_u + m^2gl^3\bar{q}_a I(2s_u + s_{ua})(1+c_a) }{ml^2(1+I^2 p_u^2)(3+2c_a)}
    .
\end{equation*}
Hence, the constrained dynamics for the acrobot under
(\ref{eqn:acrobot-constraint}) are
\begin{align}\label{eqn:acrobot-constrained-dynamics}
    \dot{q}_u &= \frac{(1+I^2 p_u^2)p_u + m^2gl^3\bar{q}_a I(2s_u + s_{ua})(1+c_a) }
            {ml^2(1+I^2 p_u^2)(3+2c_a)}
        , \\
    \dot{p}_u &= - m g l (2s_u + s_{ua})
    . \nonumber
\end{align}

@@

Letting \(I = 0\) we get the nominal system
\begin{equation}\label{eqn:khalil-perturbation-nominal}
    \begin{cases}
        \dot{x}_0 = f(t,x,0) ,\\
        x_0(t_0) = \eta_0 ,
    \end{cases}
\end{equation}
which we need to solve for the explicit solution \(x_0(t,\eta_0)\) on \([t_0,t_1]\).
One can use \(x_0(t,\eta_0)\) to compute the first-order approximation of
the nonlinear system 
\(x(t,\eta_0) = x_0(t,\eta_0) + I x_1(t,\eta_0) + R(t,I)\), where the remainder
term \(R(t,I)\) satisfies
\[
    \lim \limits_{I \to 0} \frac{R(t,I)}{I} = 0
    ,
\]
and \(x_1(t,\eta_0)\) is the solution to
\begin{equation}\label{eqn:khalil-perturbation-firstorder}
    \begin{cases}
        \dot{x}_1 = \pdiff{f}{x}(t,x_0(t,\eta_0),0)x_1 + \pdiff{f}{I}(t,x_0(t,\eta_0),0)
        , \\
        x_1(t_0,\eta_0) = 0. \\
    \end{cases}
\end{equation}

We now paraphrase Khalil's Theorem 10.1 \cite{khalil_nonlinear} on the
accuracy of perturbation analysis.
\begin{thm}
    Suppose \(f : [t_0,t_1] \times D \times [-I_0,I_0] \rightarrow \R^2\) is
    \(C^1\), and that the nominal problem (\ref{eqn:khalil-perturbation-nominal}) has a
    unique solution \(x_0(t,\eta_0) \in D\) on \([t_0,t_1]\).
    Then there exists \(I^\star > 0\) such that for all \(|I| < I^\star\), the
    solution \(x(t,I,\eta_0)\) to (\ref{eqn:khalil-setup}) satisfies
    \[
        \norm{x(t,I,\eta_0) - \left(x_0(t,\eta_0) + I x_1(t,\eta_))\right)} \leq k |I^2|
    \]
    for some \(k > 0\).
\end{thm}

The above Theorem tells us that we can approximate the solution of the nonlinear
system by the Taylor approximation 
\(x(t,\eta_0) \approx x_0(t,\eta_0) + I x_1(t,\eta_0)\).
In other words, when \(I\) is small enough solutions of the nonlinear system and
the Taylor approximation are the same up to order \(I^2\) along compact time
intervals.
Since we know that the acrobot behaves like a pendulum at \(I = 0\), we can
use this theory to more easily study the behaviour of orbits and prove energy
injection properties of our VNHC.

@@

The acrobot's constrained dynamics in
\((\alpha,\mu)\)-coordinates can be computed by evaluating
\begin{align*}
    \dot{\mu} &= \left.\pdiff{\mu(q_u,p_u)}{q}_u\dot{q_u} +
         \pdiff{\mu(q_u,p_u)}{p_u}\dot{p}_u
         \right|_{(q_u,p_u) = T\inv(\mu,\alpha)}
    , \\
    \dot{\alpha} &= \left.\pdiff{\alpha(q_u,p_u)}{q_u}\dot{q}_u + 
        \pdiff{\alpha(q_u,p_u)}{p_u}\dot{p}_u
        \right|_{(q_u,p_u) = T\inv(\mu,\alpha)}
    .
\end{align*}
with \((\dot{q}_u,\dot{p}_u)\) the given by 
 (\ref{eqn:acrobot-constrained-dynamics}).

These dynamics are much too large to write out, so we simply denote them by
\begin{align}\label{eqn:acrobot-mu-dot}
    \dot{\mu} &= f_\mu(\alpha,\mu,I)
    ,\\
    \label{eqn:acrobot-alpha-dot}
    \dot{\alpha} &= f_\alpha(\alpha,\mu,I)
    .
\end{align}
To find the nominal system we set \(I = 0\);
MATLAB's symbolic toolbox evaluates these expressions as
\begin{align}\label{eqn:acrobot-mu-dot-nom}
    \dot{\mu} &= 0
    , \\
    \label{eqn:acrobot-alpha-dot-nom}
    \dot{\alpha} &= \sqrt{\frac{6g}{5l}} 
        \sqrt{\frac{\cos(\mu\cos(\alpha)) - \cos(\mu)}{\mu^2 \sin(\alpha)^2}}
    .
\end{align}
One can confirm that \(\dot{\alpha} > 0\) for every \(\mu \in ]0,\pi[\) -- this
is expected, since the nominal system is a single pendulum.
By continuity of (\ref{eqn:acrobot-mu-dot}) with respect to \(I\), 
\(\dot{\alpha}\) remains positive for \(I\) small enough on the domain \(D\).

Suppose at time \(t = 0\) we have initialized the system at 
\((q_u,p_u) = (\mu_0,0)\), which corresponds to \((\alpha,\mu) = (0,\mu_0)\).
Since \(\alpha\) is continuously increasing, we know that the orbit will do one
of two things: it will either leave \(D\) so that the acrobot is now rotating,
or it will remain in \(D\) and hit the \(q_u\)-axis at some point 
\((\alpha,\mu) = (\pi,m)\) as is illustrated in Figure
\ref{fig:acrobot-possible-orbits}.
From there, the acrobot could escape \(D\) along the second half of the orbit or
return to some point \((\alpha,\mu) = (0,\tilde{m})\).
If \(\tilde{m} < \mu_0\) then the acrobot has lost energy; 
if \(\tilde{m} > \mu_0\) the acrobot has gained energy; 
and if \(\tilde{m} = \mu_0\), the acrobot is on a closed orbit.
Which situation occurs might depend entirely on the choice of
\(\mu_0\).

\begin{figure}
    \centering 
    % TODO: Draw the domain D, an initial point (q0,p0) = (mu0,0), and show the
    % 4 clockwise half orbits: when the orbit exits D on the bottom left corner;
    % when the orbit hits the left q-axis at (-mu0-x,0); when it hits (-mu0,0);
    % and when it hits (-mu0+x,0) for some x > 0.
    %includestandalone[width=0.5\textwdith]{images/acrobot_possible_orbits}
    \caption{If the acrobot is initialized at \((q_u,p_u) = (\mu_0,0)\), then
        either the orbit escapes \(D\) into the rotation
        zone (purple), or the orbit hits the other half of the \(q_u\)-axis 
        with a higher (red), equal (black), or lower (blue) angle than when it
        started.}
    \label{fig:acrobot-possible-orbits}
\end{figure}

Since \(\dot{\alpha} > 0\), we could theoretically find some time scaling 
\(t = \tau(\alpha)\) and use \(\alpha\) as a new ``time" variable.
Setting \(\hat{\mu}(\alpha) := \mu(\tau(\alpha))\) allows us to ignore the
dynamics of \(\alpha\) and study the evolution of \(\mu\) as a 
function of \(\alpha\) rather than as a function of \(t\) (assuming the acrobot
does not leave \(D\)).
The dynamics of \(\hat{\mu}\) are
\[
    \diff{\hat{\mu}}{\alpha} = 
    \left.\diff{\mu}{t} \diff{t}{\alpha} \right|_{t =\tau(\alpha)}
    ,
\] 
yielding the ``time"-varying scalar ODE
\begin{equation}\label{eqn:muhat-dot}
    \begin{cases}
        \diff{\hat{\mu}}{\alpha} 
        = \frac{f_\mu(\tau(\alpha),\mu,I)}{f_\alpha(\tau(\alpha),\mu,I)}
        =: g(\alpha,\mu,I)
        , \\
        \hat{\mu}(0) = \mu_0
        .
    \end{cases}
\end{equation}

In the remainder of this section we will find a first-order
approximation of \(\hat{\mu}\). 
We will show the dynamics of this approximation are \(\pi\)-periodic in
\(\alpha\), so that we only need to study whether \(\hat{\mu}\) increases or
decreases along a half-orbit \(\alpha \in [0,\pi]\).
To complete the proof of energy gain for oscillations, we will demonstrate that
\(\hat{\mu}(\pi) > \hat{\mu}(0)\).
This will imply that the the radius of orbits increases
until all orbits within \(D\) eventually leave \(D\) in finite time.

In the spirit of perturbation analysis, we expand our solution of
(\ref{eqn:muhat-dot}) into
\[
    \hat{\mu}(\alpha) = \hat{\mu}_0(\alpha) + I \hat{\mu}_1(\alpha)
    + R(\alpha,I)
    ,
\]
where (\ref{eqn:khalil-perturbation-nominal}) tells us that
\(\hat{\mu}_0\) is the solution to the nominal system
\[
\begin{cases}
    \dot{\hat{\mu}}_0 = g(\alpha,\mu,0)
    , \\
    \hat{\mu}_0(0) = \mu_0
    .
\end{cases}
\]
Since \(g(\alpha,\mu,0) = 0\) by 
(\ref{eqn:acrobot-mu-dot-nom})-(\ref{eqn:acrobot-alpha-dot-nom}), we find that
\(\hat{\mu}_0(\alpha) = \mu_0\) for all \(\alpha\).
Likewise, (\ref{eqn:khalil-perturbation-firstorder}) tells us that 
\(\hat{\mu}_1\) is the solution to
\begin{equation}\label{eqn:acrobot-mu1-dot}
    \begin{cases}
        \dot{\hat{\mu}}_1 = a(\alpha,\mu_0)\hat{\mu}_1 + b(\alpha,\mu_0)
        , \\
        \hat{\mu}_1(0) = 0
        .
    \end{cases}
\end{equation}
where
\begin{align*}
    a(\alpha,\mu_0) &= \pdiff{g}{\hat{\mu}}(\alpha,\mu_0,0)
    , \\
    b(\alpha, \mu_0) &= \pdiff{g}{I}(\alpha, \mu_0, 0)
    .
\end{align*}
These are difficult to compute by hand, so we resort to MATLAB symbolic
computation to reveal that
\begin{align*}
    a(\alpha,\mu_0) &= 0
    , \\
    b(\alpha,\mu_0) &= K \hat{b}(\alpha,\mu_0)
\end{align*}
where we define
\begin{align*}
    K &= \frac{\bar{q}_a \sqrt{30m^2g l^3}}{15}
    , \\
    \hat{b}(\alpha,\mu_0) &= \frac{
        \mu_0 |\sin(\alpha)| \left(
        5 c_{\mu_0} \cos(\mu_0 c_\alpha) - 8 \cos(\mu_0c_\alpha)^2 + 3
    \right)
    }{
    \sin(\mu_0)\sqrt{\cos(\mu_0c_\alpha) - c_{\mu_0}}
    }
    .
\end{align*}
Hence, the solution to (\ref{eqn:acrobot-mu1-dot}) is given by
\[
    \hat{\mu}_1(\alpha,\mu_0) =
    K \int \limits_0^\alpha \hat{b}(\sigma,\mu_0)d\sigma
    .
\]
